---
title: "Introduction to ggplot2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}

# More efficient file management
# install.packages("here")
library(here)

# Functions for data exploration, cleaning, and tidying
# install.packages("tidyverse")
library(tidyverse)

# Interact with Socrata open data portals
# install.packages("RSocrata")
library(RSocrata)

# Efficiently read and write data
# install.packages("readr")
library(readr)

# Graph interactivity
# install.packages("plotly")
library(plotly)

# Improved color scales
# install.packages("viridis")
library(viridis)

```

# Data Ingestion

Data was originally obtained from the City of Chicago's Open Data Portal. The query to gather the data is included, but commented out below. Data can be interacted with directly through the Socrata API endpoint using the "Socrata Query Language" (SoQL). In order to increase data ingestion speeds, you can sign up with Socrata and create an API token. Data from this query is available in the provided URL.

```{r, eval = FALSE}

# # Read in Socrata Token
# source("socrata_token.R")
# 
# # Build query using SoQL
# trip_endpoint <- "https://data.cityofchicago.org/resource/m6dm-c72p.csv"
# query_random <- "?$where=trip_id like '____%25123%25'"
# trip_endpoint_query_random <- paste0(trip_endpoint, query_random)
# 
# # Get data
# trips_random <- read.socrata(trip_endpoint_query_random,
#                              app_token = socrata_token)

```

```{r, cache = TRUE}

# Read trips data
trip_url <- "https://media.githubusercontent.com/media/rharrington31/drexel_visualization_workshop/master/trips_random.csv"
trips <- read_csv(trip_url)

```


```{r}

trips <-
  trips %>% 
  select(trip_id, trip_start_timestamp, trip_end_timestamp, trip_seconds,
         trip_miles, pickup_census_tract, dropoff_census_tract,
         pickup_community_area, dropoff_community_area, fare, tip,
         additional_charges, trip_total, shared_trip_authorized,
         trips_pooled, pickup_centroid_latitude, pickup_centroid_longitude,
         pickup_centroid_location, dropoff_centroid_latitude,
         dropoff_centroid_longitude, dropoff_centroid_location) %>% 
  mutate(shared_trip_authorized = if_else(shared_trip_authorized == "false", "Yes", "No"))

```

# When do rideshares typically start?

```{r}

trips %>% 
  ggplot(aes(x = trip_start_timestamp)) +
  geom_histogram(bins = 48)

```

# When do rideshares typically end?

```{r}

trips %>% 
  ggplot(aes(x = trip_end_timestamp)) +
  geom_histogram(bins = 48)

```

# How long do trips typically last for?

```{r}

trips %>% 
  filter(trip_seconds <= 100 * 60) %>%
  ggplot(aes(x = trip_seconds / 60)) +
  geom_histogram()

```

# How many miles is the typical trip?

```{r}

trips %>% 
  filter(trip_miles <= 100) %>% 
  ggplot(aes(x = trip_miles)) +
  geom_histogram(bins = 100)

```

# What is the relationship between the time of day that a ride begins and the time that it lasts for?

```{r}

trips %>% 
  ggplot(aes(x = trip_start_timestamp,
             y = trip_seconds)) +
  geom_point()

```

```{r}

trips %>% 
  filter(trip_seconds <= 10000) %>% 
  ggplot(aes(x = trip_start_timestamp,
             y = trip_seconds)) +
  geom_point()

```

```{r}

trips %>% 
  filter(trip_seconds <= 10000) %>% 
  ggplot(aes(x = trip_start_timestamp,
             y = trip_seconds)) +
  geom_point(alpha = 0.01)

```

```{r}

trips %>% 
  mutate(start_hour = as.numeric(format(trip_start_timestamp, format = "%H"))) %>% 
  filter(trip_seconds <= 1500) %>% 
  ggplot(aes(x = start_hour,
             y = trip_seconds / 60,
             group = start_hour)) +
  geom_boxplot()

```

# What is the relationship between trip time and distance?

```{r}

trips %>% 
  ggplot(aes(x = trip_seconds / 60,
             y = trip_miles)) +
  geom_point()

```

```{r}

trips %>% 
  ggplot(aes(x = trip_seconds / 60,
             y = trip_miles)) +
  geom_point(alpha = 0.1) +
  geom_smooth()

```

# How much do riders pay for trips?

```{r}

trips %>% 
  filter(fare <= 100) %>% 
  ggplot(aes(x = fare)) +
  geom_histogram()

```

# How much do riders generally tip?

```{r}

trips %>% 
  filter(tip <= 25, tip != 0) %>% 
  ggplot(aes(x = tip)) +
  geom_histogram(bins = 25)

```

```{r}

trips %>% 
  ggplot(aes(x = trip_miles,
             y = fare)) +
  geom_point()

```

```{r}

trips %>% 
  ggplot(aes(x = trip_seconds,
             y = fare)) +
  geom_point()

```

```{r}

trips %>% 
  ggplot(aes(x = trip_seconds,
             y = tip)) +
  geom_point()

```

# How many riders were included for a given trip?

`trips_pooled` is defined as:
> If customers were matched for a shared trip, how many trips, including this one, were pooled. All customer trips from the time the vehicle was empty until it was empty again contribute to this count, even if some customers were never present in the vehicle at the same time. Each trip making up the overall shared trip will have a separate record in this dataset, with the same value in this column.

```{r}

trips %>% 
  count(trips_pooled) %>% 
  arrange(desc(n)) %>% 
  mutate(percent = n / sum(n)) %>% 
  ggplot(aes(x = trips_pooled,
             y = percent)) +
  geom_col()

```

# How does this change if we only include trips where a shared trip was authorized?

```{r}

trips %>% 
  ggplot(aes(x = shared_trip_authorized)) +
  geom_bar()

```

```{r}

trips %>% 
  filter(shared_trip_authorized) %>% 
  count(trips_pooled) %>% 
  arrange(desc(n)) %>% 
  mutate(percent = n / sum(n)) %>% 
  ggplot(aes(x = trips_pooled,
             y = percent)) +
  geom_col()

```

```{r}

trips %>% 
  ggplot(aes(x = trips_pooled)) +
  geom_bar() +
  facet_wrap(~shared_trip_authorized)

```

# How many census tracts are represented?

```{r}

trips %>% 
  count(pickup_census_tract) %>% 
  arrange(desc(n))

trips %>% 
  count(dropoff_census_tract) %>% 
  arrange(desc(n))

```

```{r}

trips %>% 
  count(pickup_centroid_latitude, pickup_centroid_longitude, shared_trip_authorized) %>% 
  spread(key = shared_trip_authorized, value = n) %>% 
  mutate(percent_pool = `TRUE` / (`FALSE` + `TRUE`)) %>% 
  arrange(percent_pool) %>% 
  ggplot(aes(x = percent_pool)) +
  geom_histogram()

```

```{r}

trip_shares_by_centroid <-
  trips %>% 
  count(pickup_centroid_latitude, pickup_centroid_longitude, shared_trip_authorized) %>% 
  spread(key = shared_trip_authorized, value = n) %>% 
  mutate(`TRUE` = if_else(is.na(`TRUE`), as.numeric(0), as.numeric(`TRUE`)),
         `FALSE` = if_else(is.na(`FALSE`), as.numeric(0), as.numeric(`FALSE`)),
         total = `TRUE` + `FALSE`,
         percent_pool = `TRUE` / (`FALSE` + `TRUE`))

```

```{r}

pal <- colorNumeric(palette = viridis_pal()(100),
                    domain = c(0,1))

trip_shares_by_centroid %>%  
  leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap.BlackAndWhite) %>% 
  addCircleMarkers(
    lng = ~pickup_centroid_longitude,
    lat = ~pickup_centroid_latitude,
    radius = 4,
    fillColor = ~pal(trip_shares_by_centroid$percent_pool),
    stroke = FALSE,
    fillOpacity = 1,
    popup = paste0("<b>Centroid Location: </b>(", round(trip_shares_by_centroid$pickup_centroid_latitude, 3), ", ", 
                   round(trip_shares_by_centroid$pickup_centroid_longitude, 3), ")",
                   "<br><b>Rideshare Percent:</b> ", scales::percent(trip_shares_by_centroid$percent_pool),
                   "<br><b>Total Rides:</b> ", trip_shares_by_centroid$total)
  )

```

```{r}

trip_shares_by_centroid %>%  
  leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap.BlackAndWhite) %>% 
  addHeatmap(
    lng = ~pickup_centroid_longitude,
    lat = ~pickup_centroid_latitude,
    intensity = ~percent_pool,
    blur = 10,
    radius = 15,
    gradient = pal(trip_shares_by_centroid$percent_pool)
  )

```

```{r}

trips %>% 
  mutate(ID = row_number()) %>% 
  filter(str_detect(trip_id, "[0-9a-zA-Z]{4}.{0,}12.{0,}")) %>% 
  select(trip_id, ID) %>% 
  nrow() / trips %>% 
  nrow() * 45.3

```

```{r}



```

